---
title: "bm_group_project"
author: "Jin Ge"
date: "12/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(caret)
library(arsenal)
library(ggcorrplot)
theme_set(theme_minimal())
```

## load and clean the data
```{r loading_data}
law_df <- read_csv("./Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(dept = factor(dept, labels = c("Biochemistry/Molecular Biology", 
                 "Physiology", "Genetics", "Pediatrics", "Medicine", "Surgery")),
         gender = factor(gender, labels = c("female", "male")),
         clin = factor(clin, labels = c("clinical", "research")),
         cert = factor(cert, labels = c("Not certified", "Board certified")),
         rank = factor(rank, labels = c("assistant", "associate", "full")),
         dif_sal = sal95 - sal94)
```


## descrptive statistics
```{r descriptive_statistics}
## adding table labels for each variable
labels <- list(dept = "Department", 
               gender = "Gender, n(%)", 
               clin = "Clinical or Research, n(%)", 
               cert = "Board certified, n(%)",
               prate = "Publicaiton rate",
               exper = "Experience period after MD", 
               rank = "Professor title, n(%)",
               sal94 = "Salary in 1994",
               sal95 = "Salary in 1995",
               dif_sal = "Difference of salary between 1994 and 1995")

## control the descriptive information 
control_tbl <- tableby.control(
              test = F,
              total = F,
              numeric.stat = c("meansd", "medianq1q3", "iqr"),
              cat.stat = c("countpct"),
              digits = 2,
              digits.pct = 2,
              stat.labels = list(
                meansd = "Mean (SD)",
                medianq1q3 = "Median (Q1, Q3)",
                iqr = "IQR",
                countpct = "n (%)"))

tbl <- tableby(gender ~ dept + clin + cert + prate + exper + rank + sal94 + sal95 + dif_sal, data = law_df, control = control_tbl)
summary(tbl, title = "Descriptive Statistics: Salary and Covariates", labelTranslations = labels, text = T)
```

## histogram to test the normality and do transformation

```{r}
law_df %>% 
  ggplot(aes(x = sal94, y = ..density..)) + 
  geom_histogram(fill = "orange") + geom_density(alpha = .4, fill = "orange") +
  labs(x = "Salary in 1994(10^5)",
       y = "Density(10^-6)",
       title = "Histogram with density plot of Y") +
  scale_x_continuous(breaks = c(100000, 200000, 300000, 400000),
                     labels = c(10, 20, 30, 40)) +
  scale_y_continuous(breaks = c(0, 0.000002, 0.000004, 0.000006, 0.000008),
                     labels = c(0, 2, 4, 6, 8)) +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12,face = "bold")) 
## left-skewness

## do transformation of y
law_df <- law_df %>% 
  mutate(sal94 = log(sal94))

law_df %>% 
  ggplot(aes(x = sal94, y = ..density..)) + 
  geom_histogram(fill = "orange") + geom_density(alpha = .4, fill = "orange") +
  labs(x = "Natural logarithm of Salary in 1994",
       y = "Density",
       title = "Histogram with density plot of transforamed Y") +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12,face = "bold")) 
## following normal distribution
```

We neeed to use the log transformed Y to do following analysis.

## boxplot to find that something can be confounders

```{r}
law_df %>% 
  ggplot(aes(y = sal94, x = gender))+
  geom_boxplot(aes(fill = clin))+
  labs(x = 'Salary in 94')

law_df %>% 
  ggplot(aes(y = dif_sal, x = gender))+
  geom_boxplot(aes(fill = clin))+
  labs(x = 'Difference in salary between 94 and 95')
```

## automatic stepwise trying-----After consulting another senior, this step may be useless at last

```{r}
full_ml <- lm(sal94 ~ dept + gender + clin + cert + prate + exper + rank, data = law_df)
step(full_ml, direction = 'backward')
step(lm(sal94 ~ 1., data = law_df), direction = 'forward', trace = 1, scope = formula(full_ml))
step(full_ml, direction = 'both')
```

It is strange that gender is not contained in the selected model. Combined with the boxplots, there should be confounders disturbing the result.


## confounder-----most important analysis in this project
for categorical confounder, we need to create a categorical table to do independence Chi squrae test.

```{r}
## dept tested
gender <- lm(sal94 ~ gender, data = law_df)
dep <- lm(sal94 ~ gender + dept, data = law_df)
summary(gender)
summary(dep)
## >10% change of coefficient

## creating table between gender and department
dept_tl <- table(law_df$dept, law_df$gender)
dept_tl

## do chi square test to test whether two predictors are independent
chisq.test(dept_tl)
## p-value = 0.0001704 < 0.01 < 0.5. We reject the null and conclude that gender and department are associated.
## In other words, gender and dept have association meaning dept's distribution in gender has significant difference. 



## rank tested
rank <- lm(log(sal94) ~ gender + rank, data = law_df)
summary(rank)

## creating table between gender and rank
rank_tl <- table(law_df$rank, law_df$gender)
rank_tl

## do chi square test to test whether two predictors are independent
chisq.test(rank_tl)
## p-value = 0.000000003729 <<< 0.001. We reject the null and conclude that gender and rank are associated.
## In other words, gender and dept have association meaning professors' title has has a significant difference in gender. 
```

According to the Chi-square test results, department and rank are both associated with gender. Combined with the descriptive table, we find that surgery has more males than females. Full professors are more males than females. 

However, the deparment difference in gender can be seen as a relatively 'objective' condition. Females may tend to study in other fields other than surgery which has higher income. In contrast, Assigning professor titles is more likely inclined to males which can be seen as a discrimination.

Welcome to discuss since above is just a attempted interpretation.


## interaction

```{r}
interaction <- lm(log(sal94) ~ gender * clin * cert * exper, data = law_df)
summary(interaction)
```
For interaction, based on a relatively large sample size, we choose to use 0.01 as a boundary and say it is unnecessary to consider the interaction.



transformation-----log $MUST$ $DO$

$MOST$ $IMPORTANT$ 
dept~gender-----have difference but no discrimination
rank~gender-----discrimination since most female are allocated into low level of professors
We need to concerntrate on the confounder----categorical table may be $REQUIRED$

<<<<<<< HEAD
$ANOTHER$ $IMPORTANT$
vif----determine what should be deleted----'prate' can been deleted by collinearity
=======
# see if collinearity
```{r}
library(tidyverse)
library(HH)

lawsuit =  read.csv("Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    dept =factor(dept),
    gender=factor(gender),
    clin = factor(clin),
    cert = factor(cert),
    rank = factor(rank),
    lg_sal94 = log(sal94)
  )

law_origin = lawsuit %>% 
  dplyr::select(-sal94,-sal95,-id,-prate)

fit_origin = lm(lg_sal94~. ,data=law_origin)
vif(fit_origin)

fit_col1 =lm(clin~.-lg_sal94,data=law_origin)
vif(fit_col1)

fit_col2 = lm(dept~.-lg_sal94,data=law_origin)
vif(fit_col2)


```
>>>>>>> cc4a8fbe8fd8ad85b7ab67b546d790b93f7d57d1


